---
title: "lmer predictions"
author: "Cameron Roach"
date: "12/02/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load main_df
# make sure scaled variables have been created

fcst_date <- fcst_test_dates[8]
fcst_period <- 24
train_window <- 60
min_train_days <- 15

train_df <- main_df %>% 
  filter(date < fcst_date,
         date > fcst_date - train_window,
         period == fcst_period)

bid_filter <- train_df %>% 
  count(bid) %>% 
  filter(n < min_train_days) %>% 
  pull(bid)

train_df <- filter(train_df, !(bid %in% bid_filter))

test_df <- main_df %>% 
  filter(date == fcst_date,
         period == fcst_period,
         bid %in% unique(train_df$bid))  # need same levels

# Add models we wish to fit here
model_df <- list(ind_lm = fit_ind_lm(train_df),
                 ind_ns = fit_ind_ns(train_df),
                 ri = fit_ri(train_df),
                 ris = fit_ris(train_df),
                 ssc = fit_ssc(train_df)) %>% 
  map(list) %>% 
  as_tibble() %>% 
  gather(model, fit)


ssc <- model_df %>% filter(model == "ssc") %>% pull(fit) %>% .[[1]]
```



If `ssc` is one of the models, then

```{r}
pred_auto <- predict(ssc, train_df)
head(pred_auto)
```


Is the same as this

```{r}
pred_man <- getME(ssc, "Z") %*% matrix(getME(ssc, "b"), ncol = 1) + getME(ssc, "X") %*% matrix(fixef(ssc), ncol = 1)
pred_man <- as.numeric(pred_man[,1])
head(pred_man)
```

```{r}
table(as.numeric(pred_auto == pred_man))
```

Looking at the conditional mode of the random effects variable

```{r}
v <- matrix(getME(ssc, "b"), ncol = 4, byrow = TRUE)
head(v)
```

we can see that each element of the spline basis gets its own coefficient estimated.


# Variance-covariance matrices

This explains how the variance is calculated manually: [https://stats.stackexchange.com/questions/68106/understanding-the-variance-of-random-effects-in-lmer-models](https://stats.stackexchange.com/questions/68106/understanding-the-variance-of-random-effects-in-lmer-models).

```{r}
v1 <- cov(v)
# v1

# library(arm)
# U <- nrow(ranef(ssc)$bid)
# se_ranef <- se.ranef(ssc)$bid
# 1/128 * sum(se_ranef[,1]^2)  # intercept only
# v2 <- 1/U * (t(se_ranef) %*% se_ranef)


# Sum each variance-covariance matrix for each building's random effects
# Divide by number of buildings and add to the other variance component.
vars.m <- attr(ranef(ssc, condVar = TRUE)$bid, "postVar")
v2 <- 1/U * apply(vars.m, c(1, 2), sum)

v1 + v2
```

```{r}
# getME(ssc, "lambda")
v3 <- as.matrix(Matrix::bdiag(VarCorr(ssc)))
rownames(v3) <- NULL
colnames(v3) <- NULL
v3
```

```{r}
(v1+v2-v3)/v3
```

So I can use this estimated variance-covariance structure to simulate data in my other vignette. Should get simulated data that looks like the real thing.
