---
title: "SSC simulation"
author: "Cameron Roach"
date: "12 February 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(splines)
library(MASS)
library(tidyverse)
```


# Work through Durban models

## Online tutorial

Use the `sleepstudy` data set as it is similar to that in Durban's paper.

Following code is from [https://gist.github.com/tmalsburg/df66e6c2ab494fad83ee](https://gist.github.com/tmalsburg/df66e6c2ab494fad83ee).


```{r}
d <- sleepstudy
d$Subject <- factor(rep(1:18, each=10))

fm1 <- lmer(Reaction ~ Days + (Days|Subject), d)

d <- rbind(sleepstudy, sleepstudy)
d$Subject <- factor(rep(1:36, each=10))
d$Reaction <- ifelse(d$Subject %in% 19:36, NA, d$Reaction)

d$predicted <- predict (fm1, newdata=d, allow.new.levels=T)
d$simulated <- simulate(fm1, seed=1, newdata=d[-1], re.form=NA,
                        allow.new.levels=T)$sim_1

f <- function(x, ...) {
    plot(x, xlab="Subject", ylim=c(0, 500), ...)
    grid()
}
par(mfrow=c(1,3), mar=c(4,4,1,1))
with(d, f(tapply(Reaction,  Subject, mean), main="Original data", ylab="Reaction", xlim=c(1, 36)))
with(d, f(tapply(predicted, Subject, mean), main="Predicted data", ylab="", col=rep(1:2, each=18)))
with(d, f(tapply(simulated, Subject, mean), main="Simulated data", ylab="", col=rep(1:2, each=18)))
legend("bottomright", pch=c(1,1), col=1:2, c("old subjects", "new subjects"), bg="white")

x <- with(d, as.matrix(tapply(Reaction,  list(Subject, Days), mean), ncol=9))
y <- with(d, as.matrix(tapply(predicted, list(Subject, Days), mean), ncol=9))
z <- with(d, as.matrix(tapply(simulated, list(Subject, Days), mean), ncol=9))

f <- function(xlab="Day", ...) {
    plot(c(1, 10), c(0, 500), ylab="Raction", xlab=xlab, t="n", ...)
    grid()
}
par(mfrow=c(2,3), mar=c(4,4,1,1))

f(main="Original data", "")
for (i in 1:18)
  lines(x[i,])
f(main="Predicted data, old subjects", "")
for (i in 1:18)
  lines(y[i,])
f(main="Predicted data, new subjects", "")
for (i in 19:36)
  lines(y[i,], col="red")
f(main="Original data")
for (i in 1:18)
  lines(x[i,])
f(main="Simulated data, old subjects")
for (i in 1:18)
  lines(z[i,])
f(main="Simulated data, new subjects")
for (i in 19:36)
  lines(z[i,], col="red")
legend("bottomright", lty=c(1,1), col=1:2, c("old subjects", "new subjects"), bg="white")
```


## Durban random slope

This is equation (5) from Durban's paper.

First fit a linear mixed model.

```{r}
sleep.df <- sleepstudy %>% 
  rename_all(tolower) %>% 
  as_tibble()

sleep.df %>% 
  ggplot(aes(x = days, y = reaction, colour = subject)) +
  geom_line() +
  theme(legend.position = "none")

sleep.lmer <- lmer(reaction ~ days + (days|subject), sleep.df)
```

Use the variance-covariance matrix from this model to simulate data.

```{r}
m <- 100
var_cov_a <- as.matrix(Matrix::bdiag(VarCorr(sleep.lmer)))
A <- mvrnorm(m, mu = c(0, 0), 
             Sigma = var_cov_a) %>% 
  as_tibble() %>% 
  # rename(a1 = `(Intercept)`, a2 = days) %>% 
  rename_all(vars(paste0("a", 1:2))) %>% 
  mutate(subject = factor(1:m))
beta <- sleep.lmer@beta
sigma_e <- summary(sleep.lmer)$sigma

sim.df <- tibble(
  days = rep(0:9, m),
  subject = factor(rep(1:m, each=10))
) %>% 
  inner_join(A, by = "subject") %>% 
  mutate(eps = rnorm(n(), 0, sigma_e),
         reaction = beta[1] + beta[2]*days + a1 + a2*days + eps,
         data_type = "simulated")

sleep.df %>%
  mutate(data_type = "original") %>%
  bind_rows(sim.df) %>%
  ggplot(aes(x = days, y = reaction, group = subject)) +
  geom_line(size = 0.2) +
  facet_wrap(~data_type)

sim.lmer <- lmer(reaction ~ days + (days|subject), sim.df)

cat("Var-Cov matrix model fitted to original data\n")
as.matrix(Matrix::bdiag(VarCorr(sim.lmer)))
cat("Var-Cov matrix model fitted to simulated data\n")
var_cov_a
```

## Subject specific curves

Now we move into the spline section of the paper.

```{r}
sleep.df <- sleepstudy %>% 
  rename_all(tolower) %>% 
  as_tibble()

sleep.df %>% 
  ggplot(aes(x = days, y = reaction, colour = subject)) +
  geom_line() +
  theme(legend.position = "none")

#bs(sleep.df$days, df = 4, degree = 1) %>% as.data.frame() %>% dplyr::mutate(x = sleep.df$days) %>% gather(var, val, -x) %>% ggplot(aes(x = x, y = val, colour = var)) + geom_line()

K <- 2

# This includes a random intercept correlated with spline basis coefficients.
# sleep.lmer <- lmer(
#   reaction ~ days + (bs(days, df = K+1, degree = 1)|subject),
#   sleep.df
# )

# This includes an uncorrelated random intercept and correlated spline coefficients. Still not quite right.
# sleep.lmer <- lmer(
#   reaction ~ days + (1|subject) + (bs(days, df = K+1, degree = 1) - 1|subject),
#   sleep.df
# )

# We want correlated random intercept and slope, and uncorrelated random effects for spline basis coefficients where 

sleep.formula <- reaction ~ 
  days + (days | subject) +
  (bs(days, df = 2, degree = 1)[,1] - 1 | subject) +
  (bs(days, df = 2, degree = 1)[,2] - 1 | subject)
# days + (days|subject) + (bs(days, df = K+1, degree = 1) - 1 || subject),
# days + (days || subject),
# days + (bs(days, df = 2, degree = 1) || subject),

sleep.lmer <- lmer(sleep.formula, sleep.df)
sleep.lmer2 <- lmer(reaction ~ days + (days | subject) + (bs(days, df = 2, degree = 1) | subject), sleep.df)
sleep.lmer3 <- lmer(reaction ~ days + (bs(days, df = 2, degree = 1) | subject), sleep.df)
# sleep.lmer2 <- lmer(reaction ~ days + (bs(days, df = 2, degree = 1) | subject), sleep.df)

sleep.df %>%
  mutate(prediction = predict(sleep.lmer, newdata = .),
         prediction2 = predict(sleep.lmer2, newdata = .),
         prediction3 = predict(sleep.lmer3, newdata = .)) %>%
  gather(var, val, reaction, starts_with("prediction")) %>%
  ggplot(aes(x = days, y = val, colour = var)) +
  geom_line() +
  facet_wrap(~subject)

sleep.df %>%
  mutate(prediction = predict(sleep.lmer, newdata = .),
         prediction2 = predict(sleep.lmer2, newdata = .),
         prediction3 = predict(sleep.lmer3, newdata = .)) %>%
  gather(var, val, starts_with("prediction")) %>% 
  mutate(e = val - reaction) %>% 
  group_by(var) %>% 
  summarise(mean(abs(e)))
```


```{r}
m <- 10000
var_cov_a <- as.matrix(Matrix::bdiag(VarCorr(sleep.lmer)))
A <- mvrnorm(m, mu = rep(0, dim(var_cov_a)[1]), 
             Sigma = var_cov_a) %>% 
  as_tibble() %>% 
  # rename(a1 = `(Intercept)`, a2 = days) %>% 
  # rename_at(vars(paste0("a", 1:(K+2))), ~col.to) 
  rename_all(vars(paste0("a", 1:(K+2)))) %>% 
  mutate(subject = factor(1:m))
beta <- sleep.lmer@beta
sigma_e <- summary(sleep.lmer)$sigma

sim.df <- tibble(
  days = rep(0:9, m),
  subject = factor(rep(1:m, each=10))
) %>% 
  inner_join(A, by = "subject") %>% 
  mutate(eps = rnorm(n(), 0, sigma_e),
         reaction = beta[1] + beta[2]*days + a1 + a2*days + eps,
         data_type = "simulated")

# sleep.df %>%
#   mutate(data_type = "original") %>%
#   bind_rows(sim.df) %>%
#   ggplot(aes(x = days, y = reaction, group = subject)) +
#   geom_line(size = 0.2) +
#   facet_wrap(~data_type)

sim.lmer <- lmer(sleep.formula, sim.df)

cat("Var-Cov matrix model fitted to original data\n")
as.matrix(Matrix::bdiag(VarCorr(sim.lmer)))
cat("Var-Cov matrix model fitted to simulated data\n")
var_cov_a
```



# Are coefficients for splines correlated?

When knot placement can differ.

```{r}
fit.df <- sim.df %>% 
  group_by(subject) %>% 
  nest() %>% 
  mutate(fit = map(data, ~ lm(reaction ~ bs(days, 3, degree=1), .x)),
         coef = map(fit, ~ .x$coefficients))

fit.df %>% 
  mutate(coef = map(coef, ~as.data.frame(t(.x)))) %>% 
  dplyr::select(coef) %>% 
  unnest(coef) %>% 
  cor()
```


When knot placement is the same.

__TODO:__ Need to experiment with subjects that have different domains for predictor variable.

```{r}
knots_x <- quantile(sim.df$days, (1:(K+1))/(K+1))[1:K]

fit.df <- sim.df %>% 
  group_by(subject) %>% 
  nest() %>% 
  mutate(fit = map(data, ~ lm(reaction ~ bs(days, knots = knots_x, degree=1), .x)),
         coef = map(fit, ~ .x$coefficients))

fit.df %>% 
  mutate(coef = map(coef, ~as.data.frame(t(.x)))) %>% 
  dplyr::select(coef) %>% 
  unnest(coef) %>% 
  cor()
```

```{r}
simple.lmer <- lmer(reaction ~ (bs(days, 3, degree=1)|subject), sim.df)
# summary(simple.lmer)
attr(summary(simple.lmer)$varcor$subject, "correlation")
```




# Simulate electricity data

```{r}

m <- 100  # number subjects
n <- 200  # observations per subject
K <- 2  # number of knots


sigma_v <- 0.1
sigma_e <- 1
# sigma_A <- matrix(rnorm(3^2)*2-1, 3)
# sigma_A <- t(sigma_A) %*% sigma_A
sigma_A <- v3  # from lmer-predictions.Rmd
K <- ncol(sigma_A)-1
mvrnorm(1000, mu = rep(0, K+1), Sigma = sigma_A) %>% cor()

# mvrnorm(n = m,
#         mu = c(0, 0),
#         Sigma = sigma_A,
#         empirical=FALSE) %>% 
#   as_tibble() %>% 
#   ggplot(aes(x = V1, y = V2)) + 
#   geom_point()


data_list <- list()
for (i in 1:m) {
  x <- rnorm(n)
  A <- mvrnorm(1, mu = rep(0, K+1), Sigma = sigma_A)
  A <- matrix(A, ncol = 1)
  basis <- ns(x, K+1)
  beta <- rnorm(K+1, sd = sigma_v)
  
  # y <- matrix(c(rep(1, n), x, x^2, x^3), ncol = K+1) %*% A + basis %*% beta + rnorm(n, sd = sigma_e)
  y <- cbind(rep(1, n), poly(x, 3, raw = TRUE)) %*% A + 
    basis %*% beta + rnorm(n, sd = sigma_e)
  y <- as.numeric(y)
  
  data_list[[i]] <- tibble(
    subject = i,
    x = x,
    y = y,
    beta = list(beta)
  )
}

mixed_df <- bind_rows(data_list) %>% 
  mutate(subject = factor(subject))
```

```{r}
mixed_df %>%
  filter(subject %in% sample(unique(subject), 9)) %>%
  ggplot(aes(x = x, y = y, colour = subject)) +
  geom_point(shape = "o") +
  geom_smooth(se = FALSE) +
  facet_wrap(~subject) +
  theme(legend.position = "none")

mixed_df %>%
    ggplot(aes(x = x, y = y)) +
    geom_smooth(aes(colour = subject), se = FALSE, size = 0.2) +
    geom_smooth(se = FALSE) +
    theme(legend.position = "none")
```


```{r}
replicate(100, map(beta_means, ~ rnorm(1, .x)))
```



# Check model fit







